Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-03-30 docs/papers/roofline_model/gocean_roofline.tex
roofline: correct daxpy AI and update continuity-opt plot to only show sse results

@@

In order to help further understand the performance of the Continuity
kernel in Figure~\ref{FIG_cont_perf} we plot the performance of the
Continuity kernel alongside HWPC-derived metrics (obtained using the
likwid tool). We use the version compiled with SSE instructions since
that gave the best performance. This plot confirms that the $32^2$
problem size is dominated by bandwidth to L2. Sizes $64^2$, $128^2$
and $256^2$ are dominated by bandwidth to L2 and L3 and $512^2$ makes
heavy use of main memory.

\begin{figure}
  \centering
  \includegraphics[width=120mm]{nemolite2d_cont_perf}
  \caption{Performance of the Continuity kernel (open circles -- right y-axis)
    and associated HWPC data (bars -- left y-axis) as a function of
    problem size. Results are for v.16 of the Intel compiler on a Xeon
    E5-1620 v.2 CPU.}
  \label{FIG_cont_perf}
\end{figure}

As well as examining the performance of the Continuity kernel, we have
also experimented with altering the DAXPY kernel of the STREAM2
benchmark so as to vary its $AI$. We did this by adding further
arithmetic operations but making use of the same array
elements. We used five different kernels:
\begin{enumerate}
\item {\tt y(i) = y(i) + a*x(i)}, AI = 0.125
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)}, AI = 0.167
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)}, AI = 0.208
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)*y(i)}, AI = 0.25
\item {\tt y(i) = y(i) + a*x(i)+ x(i-1)*y(i)*y(i)*y(i)}, AI = 0.25
\end{enumerate}
The fifth of these introduces an un-aligned array access ({\tt
  x(i-1)}) so as to better represent real-life finite difference
kernels.  Figure~\ref{FIG_daxpy_roofline} shows the performance of
these kernels on a roofline plot for the E5-1620 CPU. The performance
of each kernel is plotted for four problem sizes, chosen so as to fit
within L1, L2, L3 cache and main memory. The performance of kernels 2,
3 and 4 is entirely consistent with the roofline model for problem
sizes that fit within the L3 and L3 caches (red and black bars). Their
performance for the smallest problem size doesn't quite reach the L1
bandwidth limit but does appear to track the roofline.  Kernel 1 has
the lowest AI of the five kernels and this appears to limit its
performance; at no point does the performance reach the limit imposed
by the bandwidth to the relevant level of the memory hierarchy. {\bf need to expore the reason for this}

>>>>

\begin{verbatim}
!DIR$ VECTOR ALIGNED
do ic = 1, nrepeat, 1
  do jj = 2, N, 1

    ! Explicit peel loop
    do ji = 2, ALIGNMENT
       rtmp1 = (sshn_u(ji  ,jj ) + hu(ji  ,jj  ))*un(ji  ,jj)
       rtmp2 = (sshn_u(ji-1,jj ) + hu(ji-1,jj  ))*un(ji-1,jj)
       rtmp3 = (sshn_v(ji ,jj )  + hv(ji  ,jj  ))*vn(ji ,jj)
       rtmp4 = (sshn_v(ji ,jj-1) + hv(ji  ,jj-1))*vn(ji,jj-1)
       ssha(ji,jj) = sshn_t(ji,jj) + (rtmp2 - rtmp1 + rtmp4 - rtmp3) * &
             rdt / area_t(ji,jj)
    end do

    do ji = ALIGNMENT+1, M, 1

       rtmp1 = (sshn_u(ji  ,jj ) + hu(ji  ,jj  ))*un(ji  ,jj)
       rtmp2 = (sshn_u(ji-1,jj ) + hu(ji-1,jj))*un(ji-1,jj)
       rtmp3 = (sshn_v(ji ,jj )  + hv(ji  ,jj  ))*vn(ji ,jj)
       rtmp4 = (sshn_v(ji ,jj-1) + hv(ji ,jj-1))*vn(ji,jj-1)

       ssha(ji,jj) = sshn_t(ji,jj) + (rtmp2 - rtmp1 + rtmp4 - rtmp3) * &
                     rdt / area_t(ji,jj)
    end do

  end do
end do
\end{verbatim}

@@

In order to help further understand the performance of the Continuity
kernel, in Figure~\ref{FIG_cont_perf} we plot the performance of the
Continuity kernel alongside HWPC-derived metrics (obtained using the
likwid tool). We use the version compiled with SSE instructions since
that gave the best performance. This plot confirms that the $32^2$
problem size is dominated by bandwidth to L2. Sizes $64^2$, $128^2$
and $256^2$ are dominated by bandwidth to L2 and L3 and $512^2$ makes
heavy use of main memory.

\begin{figure}
  \centering
  \includegraphics[width=120mm]{nemolite2d_cont_perf}
  \caption{Performance of the Continuity kernel (open circles -- right y-axis)
    and associated HWPC data (bars -- left y-axis) as a function of
    problem size. Results are for v.16 of the Intel compiler on a Xeon
    E5-1620 v.2 CPU using only the SSE SIMD instructions.}
  \label{FIG_cont_perf}
\end{figure}

As well as examining the performance of the Continuity kernel, we have
also experimented with altering the DAXPY kernel of the STREAM2
benchmark so as to vary its $AI$. We did this by adding further
arithmetic operations but making use of the same array
elements. We used five different kernels:
\begin{enumerate}
\item {\tt y(i) = y(i) + a*x(i)}, AI = 0.083
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)}, AI = 0.167
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)}, AI = 0.208
\item {\tt y(i) = y(i) + a*x(i)+ x(i)*y(i)*y(i)*y(i)}, AI = 0.25
\item {\tt y(i) = y(i) + a*x(i)+ x(i-1)*y(i)*y(i)*y(i)}, AI = 0.25
\end{enumerate}
The fifth of these introduces an un-aligned array access ({\tt
  x(i-1)}) so as to better represent real-life finite difference
kernels.  Figure~\ref{FIG_daxpy_roofline} shows the performance of
these kernels on a roofline plot for the E5-1620 CPU. The performance
of each kernel is plotted for four problem sizes, chosen so as to fit
within L1, L2, L3 cache and main memory. The performance of kernels
1--4 is entirely consistent with the roofline model for problem sizes
that fit within the L3 and L3 caches (red and black bars). Their
performance for the smallest problem size doesn't quite reach the L1
bandwidth limit but does appear to track the roofline.
